Balsa: A Compact Line Notation Based on SMILES

### 2023-01-04 manuscript.tex
Corrects conclusion.
Fixes typo.

Given an atom whose element has one or more default valences, \textit{subvalence} can be computed (Algorithm~\ref{algorithm:subvalence}). Subvalence is the number of hydrogens that can be added to an atom without exceeding the lowest possible default valence. If no suitable default valence exists, then subvalence equals zero. The algorithm begins by computing the atom's valence. Next, the ordered list of default valences of the atom's element is found. For each default valence, the difference between it and the valence is compute. If this differences is positive, it is returned as the subvalence. Otherwise, the next default valence is considered. If no suitable default valence is found, zero is returned. The subvalence of an atom whose element has no default valences (i.e., no listing in Table~\ref{table:default-valences}) is zero. For example, an iron atom always has a subvalence of zero, regardless of its bonding.

@@

This paper describes Balsa, a compact molecular serialization format designed to be forward compatible with SMILES. Constitutionally, Balsa can encode any molecule conforming to the electron-pair bonding model. In the event of undesired symmetry artifacts due to delocalization, Balsa offers a mitigation based in graph theory. Conformational isomerism of alkenes is supported by partial parity bonds. The configurations of tetracoordinate, stereogenic atoms are encoded through the use of a parity enumeration and conventions around its use.

>>>>

Given an atom whose element has one or more default valences, \textit{subvalence} can be computed (Algorithm~\ref{algorithm:subvalence}). Subvalence is the number of hydrogens that can be added to an atom without exceeding the lowest possible default valence. If no suitable default valence exists, then subvalence equals zero. The algorithm begins by computing the atom's valence. Next, the ordered list of default valences of the atom's element is found. For each default valence, the difference between it and the valence is computed. If this differences is positive, it is returned as the subvalence. Otherwise, the next default valence is considered. If no suitable default valence is found, zero is returned. The subvalence of an atom whose element has no default valences (i.e., no listing in Table~\ref{table:default-valences}) is zero. For example, an iron atom always has a subvalence of zero, regardless of its bonding.

@@

This paper describes Balsa, a compact molecular serialization format designed to be forward compatible with SMILES. Constitutionally, Balsa can encode molecules conforming to the electron-pair bonding model. In the event of undesired symmetry artifacts due to delocalization, Balsa offers a mitigation based in graph theory. Conformational isomerism of alkenes is supported by partial parity bonds. The configurations of tetracoordinate, stereogenic atoms are encoded through the use of a parity enumeration and conventions around its use.
